---
title: "SSC 442 Lab 03"
author: "Team 7"
date: "2/13/2020"
output: pdf_document
---

```{r include=FALSE}
library(tidyverse)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r include=FALSE}
rmse = function(actual, predicted) {
  sqrt(mean((actual - predicted) ^ 2))
}

get_complexity = function(model) {
  length(coef(model)) - 1
}

# https://stackoverflow.com/questions/43123462/how-to-obtain-rmse-out-of-lm-result
RMSE_2 = function(model)
{
  RSS <- c(crossprod(model$residuals))
  MSE <- RSS / length(model$residuals)
  RMSE <- sqrt(MSE)
  
  return(RMSE)
}

get_rmse = function(model, data, response) {
  rmse(actual = subset(data, select = response, drop = TRUE),
       predicted = predict(model, data))
}
```



# **Exercise 1**
```{r include=FALSE}
Ames <- read.csv("ames.csv", stringsAsFactors = FALSE)
Ames$OverallCond <- NULL
Ames$OverallQual <- NULL
Ames[is.na(Ames)] <- 0
```

```{r include=FALSE}
FullModel <- lm(SalePrice ~ ., data=Ames)
ForwardSelectStart <- lm(SalePrice ~ 1, data=Ames)

step(ForwardSelectStart, direction="forward", scope=formula(FullModel), steps=15)

comp1 <- lm(formula = SalePrice ~ Neighborhood, data = Ames)
comp2 <- lm(formula = SalePrice ~ Neighborhood + GrLivArea, data = Ames)
comp3 <- lm(formula = SalePrice ~ Neighborhood + GrLivArea + BsmtQual, data = Ames)
comp4 <- lm(formula = SalePrice ~ Neighborhood + GrLivArea + BsmtQual + 
              KitchenQual, data = Ames)
comp5 <- lm(formula = SalePrice ~ Neighborhood + GrLivArea + BsmtQual + 
              KitchenQual + RoofMatl, data = Ames)
comp6 <- lm(formula = SalePrice ~ Neighborhood + GrLivArea + BsmtQual + 
              KitchenQual + RoofMatl + BsmtFinSF1, data = Ames)
comp7 <- lm(formula = SalePrice ~ Neighborhood + GrLivArea + BsmtQual + 
              KitchenQual + RoofMatl + BsmtFinSF1 + MSSubClass, data = Ames)
comp8 <- lm(formula = SalePrice ~ Neighborhood + GrLivArea + BsmtQual + 
              KitchenQual + RoofMatl + BsmtFinSF1 + MSSubClass + BsmtExposure, data = Ames)
comp9 <- lm(formula = SalePrice ~ Neighborhood + GrLivArea + BsmtQual + 
              KitchenQual + RoofMatl + BsmtFinSF1 + MSSubClass + BsmtExposure + 
              Condition2, data = Ames)
comp10 <- lm(formula = SalePrice ~ Neighborhood + GrLivArea + BsmtQual + 
               KitchenQual + RoofMatl + BsmtFinSF1 + MSSubClass + BsmtExposure + 
               Condition2 + ExterQual, data = Ames)
comp11 <- lm(formula = SalePrice ~ Neighborhood + GrLivArea + BsmtQual + 
               KitchenQual + RoofMatl + BsmtFinSF1 + MSSubClass + BsmtExposure + 
               Condition2 + ExterQual + GarageArea, data = Ames)
comp12 <- lm(formula = SalePrice ~ Neighborhood + GrLivArea + BsmtQual + 
               KitchenQual + RoofMatl + BsmtFinSF1 + MSSubClass + BsmtExposure + 
               Condition2 + ExterQual + GarageArea + Functional, data = Ames)
comp13 <- lm(formula = SalePrice ~ Neighborhood + GrLivArea + BsmtQual + 
               KitchenQual + RoofMatl + BsmtFinSF1 + MSSubClass + BsmtExposure + 
               Condition2 + ExterQual + GarageArea + Functional + SaleCondition, data = Ames)
comp14 <- lm(formula = SalePrice ~ Neighborhood + GrLivArea + BsmtQual + 
               KitchenQual + RoofMatl + BsmtFinSF1 + MSSubClass + BsmtExposure + 
               Condition2 + ExterQual + GarageArea + Functional + SaleCondition + 
               LotArea, data = Ames)
comp15 <- lm(formula = SalePrice ~ Neighborhood + GrLivArea + BsmtQual + 
               KitchenQual + RoofMatl + BsmtFinSF1 + MSSubClass + BsmtExposure + 
               Condition2 + ExterQual + GarageArea + Functional + SaleCondition + 
               LotArea + YearBuilt, data = Ames)

Complexity <- c(1:15)

RMSEs <- c(
  RMSE_2(comp1),
  RMSE_2(comp2),
  RMSE_2(comp3),
  RMSE_2(comp4),
  RMSE_2(comp5),
  RMSE_2(comp6),
  RMSE_2(comp7),
  RMSE_2(comp8),
  RMSE_2(comp9),
  RMSE_2(comp10),
  RMSE_2(comp11),
  RMSE_2(comp12),
  RMSE_2(comp13),
  RMSE_2(comp14),
  RMSE_2(comp15)
)
```

## Part 3
```{r}
plot(
  x=Complexity,
  y=RMSEs
)
```

As complexity increases in length, the RMSE value falls exponentially. A full-size model
would not be recommended because of the sheer amount of data entries. It would be
unnecessary, as an accurate representation of the data can be achieved using much less plotted
points. This is based on the plot’s ability to capture an accurate depiction of the RMSE values
for each level of complexity.

```{r include=FALSE}
################################################################################
set.seed(9)
num_obs = nrow(Ames)

train_index = sample(num_obs, size = trunc(0.50 * num_obs))
train_data = Ames[train_index, ]
test_data = Ames[-train_index, ]

fit_0 = lm(SalePrice ~ 1, data = train_data)
get_complexity(fit_0)

get_rmse(model = fit_0, data = train_data, response = "SalePrice") # train RMSE
get_rmse(model = fit_0, data = test_data, response = "SalePrice") # test RMSE

model_list <- list(comp1, comp2, comp3, comp4, comp5)

train_rmse = sapply(model_list, get_rmse, data = train_data, response = "SalePrice")
test_rmse = sapply(model_list, get_rmse, data = test_data, response = "SalePrice")
model_complexity = sapply(model_list, get_complexity)

plot(model_complexity, train_rmse, type = "b",
     ylim = c(min(c(train_rmse, test_rmse)) - 0.02,
              max(c(train_rmse, test_rmse)) + 0.02),
     col = "dodgerblue",
     xlab = "Model Size",
     ylab = "RMSE")
lines(model_complexity, test_rmse, type = "b", col = "darkorange")
```

# **Exercise 2**
## Part 1
```{r }
model_list2 <- list(comp1, comp2, comp3, comp4, comp5, comp6, comp7, comp8, comp9, comp10,
                   comp11, comp12, comp13, comp14, comp15)

train_rmse2 = sapply(model_list2, get_rmse, data = train_data, response = "SalePrice")
test_rmse2 = sapply(model_list2, get_rmse, data = test_data, response = "SalePrice")
model_complexity2 = sapply(model_list2, get_complexity)

plot(model_complexity2, train_rmse2, type = "b",
     ylim = c(min(c(train_rmse, test_rmse)) - 0.02,
              max(c(train_rmse, test_rmse)) + 0.02),
     col = "dodgerblue",
     xlab = "Model Size",
     ylab = "RMSE")
lines(model_complexity2, test_rmse2, type = "b", col = "darkorange")
```

```{r include=FALSE}
step(ForwardSelectStart, direction="forward", scope=formula(FullModel))
forward <- lm(formula = SalePrice ~ Neighborhood + GrLivArea + BsmtQual + 
               KitchenQual + RoofMatl + BsmtFinSF1 + MSSubClass + BsmtExposure + 
               Condition2 + ExterQual + GarageArea + Functional + SaleCondition + 
               LotArea + YearBuilt + PoolQC + YearRemodAdd + Fireplaces + 
               TotalBsmtSF + BldgType + PoolArea + ScreenPorch + MSZoning + 
               BedroomAbvGr + LandSlope + LowQualFinSF + KitchenAbvGr + 
               LandContour + Condition1 + Exterior1st + MasVnrArea + Street + 
               GarageQual + GarageCond + LotConfig + TotRmsAbvGrd + WoodDeckSF + 
               GarageCars + HeatingQC + X1stFlrSF, data = Ames)

step(FullModel, direction="backward")
backward <- lm(formula = SalePrice ~ Neighborhood + GrLivArea + BsmtQual + 
                     KitchenQual + RoofMatl + BsmtFinSF1 + MSSubClass + BsmtExposure + 
                     Condition2 + ExterQual + GarageArea + Functional + SaleCondition + 
                     LotArea + YearBuilt + PoolQC + YearRemodAdd + Fireplaces + 
                     TotalBsmtSF + BldgType + PoolArea + ScreenPorch + MSZoning + 
                     BedroomAbvGr + LandSlope + LowQualFinSF + KitchenAbvGr + 
                     LandContour + Condition1 + Exterior1st + MasVnrArea + Street + 
                     GarageQual + GarageCond + LotConfig + TotRmsAbvGrd + WoodDeckSF + 
                     GarageCars + HeatingQC + X1stFlrSF, data = Ames)

# compBackward according to the step model is a slightly better fit, so I will test the RMSEs of
# the backward selection model

train_rmse_backward_selection = get_rmse(backward, train_data, "SalePrice")
test_rmse_backward_selection = get_rmse(backward, test_data, "SalePrice")
model_complexity_backward_selection = get_complexity(backward)

```

## Part 3

Our predicted sale price was 29513.22. This was determined using important
factors such as lot area and size, exterior quality, condition of the home, basement finish,
location, and other key components in predicting sale price. Generally, factors like these
are taken into account when selling a house, therefore, our prediction should be mostly
accurate. We are confident that our group’s prediction will be among the lowest because
of how close our value (29513.22) is compared to the minimum RMSE value of 25000.